---
title: "Lab 1"
author: "Mario Giacomazzo"
date: "August 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21)

**Assigned:** Friday, August 24, 2018

**Due:** Friday, August 31, 2018 by 5:00 PM

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a html file generated in RStudio. All assigned problems are chosen according to the online textbook (http://r4ds.had.co.nz/). 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense?

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
```{r}
# What's gone wrong with this code? Why are the points not blue?

# --> Original graph
ggplot(data=mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue") 
  ) +
  ggtitle("Screwed up Graph")

# --> Fixed graph
ggplot(data=mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), color = "blue"
  ) +
  ggtitle("Fixed Graph")

#--> Solution:
# The color property must be set outside of aes.

```

##  Exercise 2
```{r}
# Which variables in mpg are categorical?
?mpg

#--> Categorical variables:
# 1) manufacturer
# 2) model
# 3) trans
# 4) drv
# 5) fl
# 6) class

#--> Check by looking at variables:
head(mpg)

#--> Conclusion:
# The six variables listed above are the categorical variables.

```

##  Exercise 5
```{r}
# What does the storke aesthetic do? What shape does it work with?

#--> Familiarize with geom_point
?geom_point

#--> Solution:
# Stroke controls the border of shapes that have a border (e.g. circle, square, triangle, diamond). Evidently--based on the snippet below--this is every one of the 25 shapes in R. One can control the fill as well, which is the thickness of the center.

#--> Graph a sample graph with each of the 20 shapes:
for(i in 0:25){
  print(ggplot(data=mpg) +
        geom_point(
        mapping = aes(x = displ, y = hwy), color = "blue",  shape = i, stroke = 4
        )
  )
}
```

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 4
```{r}
#
```

##  Exercise 5

I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.


## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 
