---
title: "Lab 1"
author: "Mario Giacomazzo"
date: "August 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21)

**Assigned:** Friday, August 24, 2018

**Due:** Friday, August 31, 2018 by 5:00 PM

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a html file generated in RStudio. All assigned problems are chosen according to the online textbook (http://r4ds.had.co.nz/). 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense?

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive  
- "r" = rear-wheel drive  
- "4" = 4-wheel drive  

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1

What's gone wrong with this code? Why are the points not blue?  
The color property must be set outside of aes (see code below).

```{r}

# --> Original graph
ggplot(data=mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy, color = "blue") 
  ) +
  ggtitle("Screwed up Graph")

# --> Fixed graph
ggplot(data=mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy), color = "blue"
  ) +
  ggtitle("Fixed Graph")

```

##  Exercise 2
Which variables in mpg are categorical?

Categorical variables:
  - manufacturer
  - model
  - trans
  - drv
  - fl
  - class
 
```{r}
?mpg

#--> Check by looking at variables:
head(mpg)

#--> Conclusion:
# The six variables listed above are the categorical variables.

```

##  Exercise 5
What does the stroke aesthetic do? What shape does it work with?  

Stroke controls the border of shapes that have a border (e.g. circle, square, triangle, diamond). Evidently--based on the snippet below--this is every one of the 25 shapes in R. One can control the fill as well, which is the thickness of the center.
```{r}

#--> Familiarize with geom_point
?geom_point

#--> Graph a sample graph with some of the 20 shapes:
for(i in 0:2){
  print(ggplot(data=mpg) +
        geom_point(
        mapping = aes(x = displ, y = hwy), color = "blue",  shape = i, stroke = 4
        )
  )
}
```

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?  

  - Line chart --> line geom (geom_line)
  - Boxplot --> boxplot geom (geom_boxplot)
  - Histogram --> histogram geom (geom_histogram)
  - Area --> area geom (geom_area)

##  Exercise 2
Run this code in your head and predict what the output will look like.  

I predict that this will be a scatter plot with a smoothed line running through it. The standard error (gray shading around the line) is turned off. On the x-axis is the engine displacement. On the y-axis is highway mileage. The drivetrain variable is shown with color.

```{r}
# Graph from ex. 2
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = drv)
  ) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
  

I missed that the smoothed lines would be seperated based on the drivetrain--it's pretty clear now though!

##  Exercise 3

What does show.legend =  FALSE do? What happens if you remove it? Why do you think I used it earlier in this chapter?  

  - show.legend = FALSE turns off the legend. See the graph below. It is the same as from page 8 of the text book, but without a legend.
  - If you remove it (I'm taking "it" to mean "show.legend = FALSE") then the legend will appear again.
  - You removed it earlier in the chapter to reduce clutter--there were three faceted figures side-by-side and having a legend would have been a bit too much.

```{r}
#--> Graph from page 8 sans legend
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class), show.legend = FALSE)

```

##  Exercise 4

What does the se argument to geom_smooth() do?  

se displays the confidence interval around the fitted curve. Solution courtesy of <https://ggplot2.tidyverse.org/reference/geom_smooth.html>

```{r}
#--> Example graph with se=FALSE
ggplot(data = mpg,
       mapping = aes(x = displ, y = cty)
 ) + 
  geom_point() +
  geom_smooth(se = FALSE, color = "red")
```

##  Exercise 5
(Solution provided by Dr. Mario below)  
I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.

```{r}
#--> Get an idea of what's in the dataset
head(diamonds)

```

Categorical:   
  - cut
  - color
  - clarity
  
Numerical:   
  -carat
  -depth
  -table
  -price
  -x
  -y
  -z
  
Now that we know what's in the data set, we can start poking around. We will most likely be interested in how all of the variables factor into the price of the diamond. Price will be our numeric; we'll try cut and color as categorical variables.

```{r}

#--> Graph relating price, cut, and color
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = color, y = price, color = color)) + 
  facet_wrap(~ cut) +
  ggtitle("Cut and color relationship with price")

#--> Note:
# Turns out this is very nearly identical to Dr. Mario's. Did not intend for that to happen. For good measure, here is another visualization that is different from Dr. Mario's.
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = clarity, y = price, color = clarity)) + 
  facet_wrap(~ cut) + 
  theme(axis.text.x = element_text(angle=60, hjust=1)) + # Make the x labels so they are not running into each other
  ggtitle("Cut and clarity relationship with price")
  
```


## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 
 
 We're still interested in price, color, and cut. Let's also add in carat.
 
```{r}

#--> Imagining point plot object relating carat and price; color coded by color and faceted by cut.
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price, color = color), size = 0.6) +
  facet_wrap(~ cut) + 
  ggtitle("Diamond cut, carat, and color relationship with price") + 
  xlab("Carat") +
  ylab("Price (USD)")
  
```
 
 This lab completed by Coleman Breen.
